\subsection{Tool installation}
\subsubsection{Installation of \IonTorrent~mapping software \texttt{tmap}}
To detect variants present in the data we need to map sequencing
reads against reference sequence. For reads generated with \IonTorrent~
sequencing platform we will use program \texttt{tmap}.
%Different short read mappers perform
%with various success depending on sequencing platform. 

\texttt{tmap} and its installation instructions can
be found at \url{https://github.com/iontorrent/TS/tree/master/Analysis/TMAP}.
Since we are do not have the administrator's rights on this server, we can't install
software on the server. However, we can still 
Compilation instructions:\\~\\
%https://github.com/iontorrent/TS/tree/master/Analysis/TMAP
\texttt{cd \textasciitilde/\progDir}\\
\texttt{git clone git://github.com/iontorrent/TMAP.git}\\
\texttt{cd TMAP}\\
\texttt{git submodule init}\\
\texttt{git submodule update}\\
\texttt{sh autogen.sh}\\
\texttt{./configure}\\
\texttt{make}\\

Lets test the program to confirm that it was compiled successfully:\\~\\
\texttt{./tmap}\\

If you see the programs interface, the program was installed successfully.
Move the compiled binary file to our \texttt{\binDir}~folder:\\~\\
\texttt{cp tmap \textasciitilde/\binDir}\\
%cp ~/data/samtools. .
%(unzip)
%unzip samtools.zip
%cd samtools-develop
%(make)
%(zlib1g-dev)
%(libncurses5-dev)
\subsubsection{Installation of \texttt{samtools} and \texttt{bcftools}}
We will also need two tools named \texttt{samtools} and \texttt{bcftools} which are used for
manipulation of mapped reads and variation calling. You can obtain these tools from
\url{http://sourceforge.net/projects/samtools/}. Click on \texttt{Files} $\rightarrow$ \texttt{samtools} $\rightarrow$ \texttt{1.2}
Rightclick on \texttt{samtools-1.2.tar.bz2} and choose \texttt{Copy link address}. 
We will download these tools in directory \texttt{\progDir}:\\~\\
\texttt{cd \textasciitilde/\progDir}\\
\texttt{wget -O samtools.tar.bz2}\\
paste the copied location and hit \texttt{Enter}.

Now, repeat this process for \texttt{bcftools}:\\~\\
\texttt{wget -O bcftools.tar.bz2}\\
paste the copied location and hit \texttt{Enter}.

The tools are compressed in \texttt{.tar.bz2}
format, so we need to extract them:\\~\\
\texttt{tar -jxvf samtools.tar.bz2}\\
\texttt{tar -jxvf bcftools.tar.bz2}\\

\begin{framed}
Note that if we had a lot more files to extract and it would be too
time consuming to manually extract them, we could use a \texttt{for}
loop and pattern matching to extract archives automatically:\\~\\
\texttt{for archive in *.tar.bz2; do}\\
\texttt{\indent tar -jxvf \$archive}\\
\texttt{done}\\
\end{framed}
%We will download its
%source code and library that it depends on with \git.
%To to download \texttt{samtools} and its library \texttt{htslib}, type:\\~\\
%\texttt{cd \progDir}\\
%\texttt{git clone https://github.com/samtools/samtools}\\
%\texttt{git clone https://github.com/samtools/htslib}\\

We have downloaded and extracted source code of the tools, but to make
these tools usable, we need to compile the source code. Source code compiling
is performed with command \texttt{make}:\\~\\
\texttt{cd samtools-1.2}\\
\texttt{make}\\
%\texttt{cd ../../bcftools-1.2}\\

If (hopefully) no errors were encountered, then \texttt{samtools} was compiled correctly. type:\\~\\
\texttt{./samtools}\\

to test the tool. If you see the program's interface, then the program was compiled successfully.
Move compiled binary file to a directory where we are storing our compiled software:\\~\\
\texttt{cp samtools \textasciitilde/\binDir\\}

Repeat the same process for \texttt{bcftools}
(go to \texttt{bcftools} source code directory,
compile it and copy resulting binary file to \texttt{\textasciitilde/\binDir})
